---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# gblidar

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->

# To Do List:

- [x]  Added options for merging rasters and exporting as SpatRaster, stars of character.
- [x]  Generic merge for all rasters in gbl_catalog.
- [x]  Add special functions for hitting the WCS data to get faster access to large composite areas.
- [ ]  Add Scottish Data
- [ ]  Add Welsh Data
- [ ]  Write Tests

**Scottish and Welsh Data are yet to be added as the tricky thing to get right 
is the English ESRI API stuff - once this is stable we can slot the other nations
in with relative ease I hope**


## Installation

You can install the development version of gblidar from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pkg_install("h-a-graham/gblidar")
```

## Example

This is a super quick demo where we can query the Environment Agency's entire
LIDAR catalogue using the `eng_search` function. This will return a `gbl_catalog`
object which can be used to filter the assets and then merge them into a single
raster. In this example we search the area around Scafell Pike, use the 
`filter_catalog` function to reduce the number of assets to just the 2m DTM 
composite for 2022, and then use the `merge_assets` function to merge all the 
assets into a single raster. By default the raster is saved to disk and the   
file path is returned, but this can be changed using `raster_class` argument 
in `merge_assets` to return a `stars` or `SpatRaster` object (or you can 
alternatively set the global option `gblidar.out_raster_type` to "stars" or 
"SpatRaster" as in the example below).


```{r example}
library(gblidar)
library(sf)
if (rlang::is_installed("terra")) {
  library(terra)
  options(gblidar.out_raster_type = "SpatRaster")
}

scafell_box <- st_point(c(321633, 507181)) |>
  st_buffer(2000) |>
  st_sfc() |>
  st_set_crs(27700)

scafell_catalog <- eng_search(scafell_box)

print(scafell_catalog)

DTM_catalog <- scafell_catalog |>
  filter_catalog(
    product == "LIDAR Composite DTM",
    resolution == 2,
    year == 2022
  )

print(DTM_catalog)

scafell_raster <- merge_assets(DTM_catalog, mask = TRUE)

plot(scafell_raster, col = grDevices::hcl.colors(50, palette = "Sunset"))
```

The EA have now released their latest composite products as WCS services 
(Awesome!). To make best use of this, we provide the `eng_composite` function
which can be used to directly download only the data you need for any area 
in England. This is significantly faster than the `eng_search` function and is
preferred if you are only interested in the latest composite products. Both
elevation and hillshade data are available for "fz_dsm" (first return DSM),
"dsm", "dtm", and "vom". 

```{r example2}
options(gblidar.progress = FALSE) # for readability in this example.

search_box <- st_point(c(532054, 181145)) |>
  st_buffer(500) |>
  st_sfc() |>
  st_set_crs(27700)

# the elevation data for the first return DSM
fz_dsm <- eng_composite(search_box, product = "fz_dsm")
# the hillshade data for the last return DSM
dsm_hs <- eng_composite(search_box, product = "dsm", product_type = "hillshade")

par(mfrow = c(1, 2))
plot(fz_dsm, col = hcl.colors(150, "mako"))
plot(dsm_hs, col = hcl.colors(256, "Blues"), legend = FALSE)
```